### 2016-08-05 1607 - JMC for g2/Paper/JMC_dis.tex
Added ETL sum rule discussion. Revised paper from beginning until conclusion, excluded; updated e+e- diagrams with right notation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% On jet mass corrections and transversity
%% May 10, 2013
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[preprintnumbers,floatfix,nofootinbib]{revtex4}

\usepackage[tbtags]{amsmath}  % AMS math 
\usepackage{amssymb}          % AMS symbols
\usepackage{bm}               % bold math
\usepackage{graphicx}         % PostScript figures
\usepackage[export]{adjustbox}% to align images
\usepackage{hhline,multirow}  % for nicer tables
\usepackage{dcolumn}          % Align table columns on decimal point
\usepackage{slashed}
\usepackage{datetime}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

%%%%%% for draft
\newcommand{\todo}[1]{\marginpar{$\bullet$}\textbf{#1}}
\def\AAcom#1{{\bf  \textcolor{Red}{[AA: {#1}]}}}
\def\AAmod#1{{\textcolor{Green}{#1}}}

%%%%%%  Definitions   %%%%%%%%%%%

\newcommand{\Pslash}{P \hspace{-0.24cm} / \,}
\newcommand{\kslash}{k \hspace{-0.21cm} / \,}
\newcommand{\lslash}{l \hspace{-0.19cm} / \,}
\newcommand{\nslash}{n \hspace{-0.24cm} / \,}
\newcommand{\vslash}{v \hspace{-0.21cm} / \,}
\newcommand{\Sslash}{S \hspace{-0.22cm} / \,}
\newcommand{\Dslash}{D \hspace{-0.22cm} / \,}
\newcommand{\epsslash}{\varepsilon \hspace{-0.18cm} / \,}
\newcommand{\Tr}{\operatorname*{Tr}\nolimits} % Trace operator
\newcommand{\xbj}{{x_B}}                   % Bjorken variable
\newcommand{\de}{d}                    % differential
\newcommand{\ii}{i}                    % imaginary unit

\newcommand{\eg}{{\em e.g.}}
%\newcommand{\mj}{\langle m_j \rangle}
%\newcommand{\mj}{M_q^\text{jet}}
\newcommand{\mj}{M_q}
\newcommand{\mq}{m_q}
\newcommand{\mjs}{\langle m_j^2 \rangle>}

\allowdisplaybreaks[2]

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%   TITLE    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\preprint{JLAB-THY-16-????}

\title{Accessing the nucleon tensor structure in inclusive deep inelastic scattering}

\author{Alberto~Accardi$^{a}$, Alessandro~Bacchetta$^{b}$} 
\affiliation{
$^a$Hampton University, Hampton, VA 23668, USA,
and Jefferson Lab, Newport News, VA 23606, USA \\
$^b$Dipartimento di Fisica, Universit\`a degli Studi di Pavia, and INFN,
Sez. di Pavia, 27100 Pavia, Italy
}

\date{\bf \today, \currenttime}

\begin{abstract}
We revisit the standard analysis of inclusive Deep Inelastic Scattering (DIS)
off nucleons taking into
account the fact that on-shell quarks cannot be present in the final state,
but they rather decay into hadrons. As a consequence, a spin-flip term
associated with the invariant mass of the produced hadrons is generated
non perturbatively, and couples to the target's transversity distribution
function. In inclusive cross sections, this provides an hitherto neglected and
large contribution to the twist-3 part of the $g_2$ structure function, that
can explain the discrepancy between recent calculations and fits of this
quantity. It also provides an extension of both the Burkardt-Cottingham sum rule, putting constraints on the small-$x$ behavior of the transversity
function and can help constraining the nucleon tensor charge, and of the Efremov-Teryaev-Leader sum rule, suggesting a novel way to mwasure the tensor charge of the proton. 
%Perspectives to measure the new spin-flip term, and applications to
%spin-1 targets will be briefly discussed. 
\end{abstract}

%\pacs{ }

%\keywords{ }

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The tensor charge is a fundamental property of the nucleon, that is at
present poorly constrained but of fundamental importance, not the least because its knowledge can be used also to put constraints on the search of physics beyond the Standard Model~\cite{Cirigliano:2013xha,Bhattacharya:2015esa,Courtoy:2015haa}.
The tensor charge has been estimated in lattice QCD (see, \eg,
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn}), but
only limited information is available from direct measurements. Its experimental extraction requires first of all the measurement of the so-called transversity parton distribution function,
denoted by $h_1^q(x)$ (see Ref.~\cite{Barone:2001sp} for a review on transversity and
Refs.~\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa} for the most recent
extractions). Secondly, one needs to perform a flavor decomposition and evaluate its integral, that corresponds to the contribution of flavor $q$ to the tensor charge.

The transversity distribution is notoriously difficult to access because it is
a chiral-odd function and needs to be combined with a spin-flip mechanism to
appear in a scattering process~\cite{Jaffe:1996zw}. Usually, this spin flip is provided by another
nonperturbative distribution or fragmentation function, accessible in Drell-Yan or semi-inclusive DIS~\cite{Ralston:1979ys,Jaffe:1991kp,Jaffe:1993xb,Collins:1992kk}.
The only other known way to attain spin-flip terms in QED and QCD is taking into
account mass corrections. In fact, it is well known that the transversity distribution gives a
contribution to the structure function $g_2$ in inclusive DIS (see, \eg,
\cite{Accardi:2009au} and references therein), and in
particular to the violation of the so-called Wandzura--Wilczek relation for
$g_2$~\cite{Wandzura:1977qf}. However, this contribution is proportional to the current quark mass and can be expected to be negligibly small. This implies that
transversity cannot be accessed in inclusive DIS, but only in more complex processes such as semi-inclusive DIS or dilepton production in hadron-hadron collisions.

In this paper, we discuss a novel way of accessing the transversity parton distribution function (PDF) and measuring the proton's tensor charge in totally inclusive Deep Inelastic Scattering. 
We revisit the standard analysis of inclusive DIS taking into account the fact
that on-shell quarks cannot be present in the final state, but they rather
decay into hadrons (ideally, forming jets of hadrons). This is sufficient to modify the structure of the DIS
cut-diagram, even if none of the hadrons is detected in the final
state. For a proper description of this effect, we include ``jet
correlators'' into the analysis, and pay particular attention to ensuring that our results are gauge invariant. We observe that the inclusion of jet correlators introduces a new
contribution to the inclusive $g_2$ structure function. This term is
proportional to the transversity distribution function multiplied by a
nonperturbative ``jet mass'' parameter, likely much larger than the mass of light quarks, and has the
interesting features that: (a) it violates the Wandzura--Wilczek relation; (b)
it extends the Burkhardt--Cottingham sum rule, providing a useful bound on the small-$x$ behavior of the transversity distribution; (c) it also extends the Efremov-Teryaev-Leader sum rule, and provides a novel way to measure the proton's tensor charge.
We provide estimates of this new jet-mass induced contribution based on a recent extraction of transversity, and show that it can indeed be very large.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Jet correlator and twist-2 structure functions}
\section{The jet correlator}

Motivated by large-$x$ mass corrections to inclusive DIS structure functions,
Accardi and Qiu have introduced in the LO handbag diagram 
a ``jet correlator'' (also called ``jet factor'' by Collins and Rogers in Ref.~\cite{Collins:2007ph})
that accounts for invariant mass production in the current jet, and ensures
that leading twist calculations in collinear factorization are consistent with
the requirement imposed by baryon number conservation that $x_B<1$
\cite{Accardi:2008ne}. The jet correlator is depicted in
Figure~\ref{fig:handbags}(a) and is defined as 
\begin{align} 
\Xi_{ij}(l,n_+) =\int
  \frac{\de^4\eta}{(2\pi)^4}\; e^{i l \cdot \eta}\,
    \langle 0|\, {\cal U}^{n_+}_{(+\infty,\eta)}
\,\psi_i(\eta)
             \bar{\psi}_j(0)\,
{\cal U}^{n_+}_{(0,+\infty)}\,   |0 \rangle \ ,
\label{e:xifull}
\end{align} 
In this definition, $l$ is the quark's four-momentum, $\Psi$ the quark field
operator (with quark flavor index omitted for simplicity), and $|0\rangle$ is
the nonperturbative vacuum state. Furthermore, we explicitly guarantee the
correlator's gauge invariance by introducing two Wilson line operators ${\cal
  U}^{n_+}$ along a light-cone plus direction determined by the vector
$n_+$. This path choice for the Wilson line is required by QCD factorization
theorems, and the vector is determined by the particular hard process to which
the jet correlator contributes. For example, in the case of inclusive DIS
discussed in this paper, this is determined by the four momentum transfer $q$
and the proton's momentum $p$.

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions (PCFs), using the vectors $l$ and $n_+$: 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ ,
\label{e:jetexpansion}
\end{equation} 
where $\Lambda$ is an arbitrary scale, introduced for power counting purposes.
In this parametrization, no terms proportional to $\gamma_5$ enters because of parity invariance. Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes only at twist-4 order, and will not be considered further in this paper. We focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
    + O(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the cut quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one (or, in other words, the interacting rather than the free quark fields).
The $A_1$ and $A_2$ terms can be nicely interpreted in terms of the spectral
representation of the cut quark propagator (see, e.g., Sec.~6.3 of
\cite{D'Hoker:2004aa} and Sec.~2.7.2 of \cite{Romao:2013aa}),
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu_j^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ is interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a), and the $J_i$ are the spectral functions of the
quark propagator, also called ``jet functions'' in
\cite{Accardi:2008ne}. These satisfy
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqns.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

\begin{figure*}[bt]
  \centering
  (a)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram0}
  \hfill
  (b)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram2}
  \hfill
  (c)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram1}
  \caption{Diagrams contributing to DIS scattering up to twist-3, including a jet correlator in the top part. Note the gluon attaches to both the nucleon and jet correlators. The Hermitian conjugates of diagrams (b) and (c), i.e., with gluons attaching to the right of the cut, are not shown.
  }
  \label{fig:handbags}
\end{figure*}

When inserting the jet correlator in the handbag diagram for inclusive DIS,
the invariant jet mass $\mu^2$ is integrated from 0 to $Q^2(1/x_B-1)$, where $x_B$ is the Bjorken invariant. This
induces (kinematical) corrections of order $O(1/Q^2)$, whose effect on the
$F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}: 
\begin{align}
  F_2(x_B) & = \int_0^{Q^2(1/x_B-1)}d\mu^2\, J_2(\mu^2) F_2^{(0)}(\xi(1+\mu^2/Q^2)) \ ,
\label{eq:F2}
\end{align}
where $F_2^{(0)}$ is the structure function calculated with the handbag
diagram sporting a bare quark propagator instead of the jet correlator, and
$\xi=2x_B/(1+\sqrt{1+4x_B^2M^2/Q^2})$ with $M$ the nucleon's mass is the
Nachtmann scaling variable. (We omitted the dependence of the structure
function on $Q^2$ for clarity of notation). In this paper we limit our
attention to effects of order $O(1/Q)$ and therefore can extend the
integration to $\mu^2=\infty$. As a consequence, the jet function $J_2$ decouples and, thanks to the sum rule \eqref{eq:jetfnsprops}, integrates to 1. One then recovers the conventional result, 
\begin{align}
  F_2(x_B) = \Big[ \int_0^\infty d\mu^2\, J_2(\mu^2) \Big] F_2^{(0)}(x_B) 
     + O(\Lambda^2/Q^2) = F_2^{(0)}(x_B)  + O(\Lambda^2/Q^2) \ .
\end{align}
%The same holds true also for the helicity structure function $g_1$.
%
More in general, the jet correlator decouples from the parton correlator $\Phi$ in any inclusive cross section calculation up to $O(1/Q)$, and the inclusive structure functions only depend on the integrated jet correlator 
\begin{equation} 
  \Xi(l^-) \equiv \int \frac{dl^2}{2l^-} d^2 l_T \, \Xi(l) 
    =  \frac{\Lambda}{2 l^-}\,\xi_1 {\bm 1}
    +  \xi_2 \frac{\nslash_-}{2} 
    + \text{higher\ twists}
    % +\frac{\Lambda^2}{4 (l^-)^2}\,\xi_3 \nslash_+
    % + i \frac{\Lambda}{2 l^-} \xi_4 
    % \frac{ \bigl[\nslash_-, \nslash_+ \bigr]}{2}.
\end{equation} 
where 
\begin{align}
\xi_1 &= \int d\mu^2 \frac{\mu}{\Lambda} J_1(\mu^2) 
       \equiv \frac{\mj}{\Lambda},
&
\xi_2 &= \int d\mu^2 J_2(\mu^2) = 1 \ .
% \\
% \xi_3 &= 0,
% &
% \xi_4 &= \int d\mu^2 \frac{\mu^2}{\Lambda^2} J_1(\mu^2) = \frac{\mjs}{\Lambda^2}.
\end{align} 
Here, $\mj$ can be interpreted as the average invariant mass produced in the spin-flip fragmentation processes of a quark of flavor $q$.
% The $\xi_1$ and $\xi_4$ have an interpretation as average invariant mass (and mass squared) of the current jet, and $l^-=q^- + O(1/Q)$ is imposed by 4-momentum conservation at the hard scattering vertex. The $\xi^4$ term contributes at order $O(1/Q^2)$ and is no further considered in this paper. 
It is important to notice that $\xi_2=1$ exactly due to CPT invariance
(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), while $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is
dynamically determined. From the analytic properties of spectral functions we
may expect \cite{Accardi:2008ne} $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2
(\mu^2) \theta (\mu^2-m_\pi^2)$ with the continuum starting at $m_\pi$, the
mass of the pion, due to color confinement effects. Taking into account that
$J_1 < J_2$, we may therefore expect  
\begin{align}
  \label{eq:mjet}
  \mj = O(100 \text{ MeV}) \ .
\end{align}
Although $\mj$ is in general a nonperturbative quantity, it is interesting to
notice that  
\begin{align}
  \label{eq:xi2_chiral_cond}
  \mj = \frac{\Lambda}{4} \int \Tr \big[ \Xi(l) \bm 1] 
   = \langle 0 | \bar \psi_i(0) \psi_i(0) | 0 \rangle
\end{align}
% \todo{[AA] 2 issues here: (1) possible factors of 4. 
% (2) I am neglecting the sum over quarks; we have to fix that somehow but there is a little tension between the quark level $\xi^a$ and the contribution to structure function which is summed and weighted by the electron charge.} 
On the non-perturbative vacuum, this corresponds the contribution of a quark of flavor $q$ to the chiral condensate, and is calculable in principle in lattice QCD.
Considering the perturbative vacuum and calculating the same to leading order in the strong coupling constant, corresponds, instead, to taking the trace of the cut bare-quark propagator to obtain 
$\mj = \,_{\text{pert}} \langle 0
|  \bar \psi_i(0) \psi_i(0) | 0 \rangle_{\text{pert}}  = \mq$, with $\mq$ the quark mass. One recovers, therefore, the result of the calculation with the conventional handbag diagram. 
However, we are here considering non perturbative effects
on the quark fragmentation and $\mj \gtrsim \mq$. Therefore, differently from $J_2$, the $J_1$ function leaves an imprint on the inclusive DIS cross section even in the asymptotic $Q^2 \to \infty$ regime.

%, an example of which we will discuss in the next section.  
%very importantly, this
%quantity can be computed in lattice QCD. 
%\todo{[AA] Are qe sure that $\mj > \mq$??}
%\todo{[AA] Need to discuss connection to quark condensate and tensor charge.}

\section{Twist-3 analysis}

Extending the analysis of \cite{Accardi:2008ne} to the calculation of twist-3
structure functions requires not only to consider the $\xi_1$ term in the jet
correlator, but also quark-gluon-quark correlators in both the proton and the
vacuum as depicted in Figs.\ref{fig:handbags}(b) and (c), respectively. 
In the former the $\xi_1$ terms contribute to $O(1/Q^2)$, so that up to $O(1/Q)$
these give the same contribution as in the conventional handbag calculation.

The novel element in our analysis is the jet's quark-gluon-quark correlator
$\Xi_A^{\mu}(l,k)$ in diagram \ref{fig:handbags}(c), 
\begin{equation} 
\begin{split} 
  \left(\Xi_A^{\mu} \right)_{ij} &=
   \frac{1}{2}\, \sum_X \int \frac{\de \eta^+\, \de^2 \bm{\eta}_T}{(2\pi)^{3}}\;
   e^{\ii k \cdot \eta}\,
   \langle 0|\,
   {\cal U}^{n_+}_{(+\infty,\eta)}\,
   g A^{\mu}(\eta)\,
   \,\psi_i(\eta)|X\rangle
   \langle X|
             \bar{\psi}_j(0)\,
   {\cal U}^{n_+}_{(0,+\infty)}
   |0\rangle \bigg|_{\eta^-=0} .
\label{e:xi_A}
\end{split} 
\end{equation}  
This diagram and its Hermitian conjugate are not only important to account for
all contribution of order $O(1/Q)$, but also to restore gauge invariance, which is broken in diagram \ref{fig:handbags}(a) due to the different mass
of the incoming and outgoing quark lines, namely, $\mq \neq \mj$.

Rather than directly using the definition \eqref{e:xi_A}, it is convenient to
calculate the inclusive cross section as an integral of the semi-inclusive
one, utilize the QCD equation of motions, sum over all
hadron flavors, and take advantage of  
\begin{align}
  \label{eq:SIDIS_to_DIS}
  \sum_h \int \frac{d^3p_h}{(2\pi)2E_h} \Delta^h(l,p_h) = \Xi(l) \ , 
\end{align}
where $\Delta^h$ is the quark fragmentation correlator for production of a
hadron of flavor $h$ and momentum $p_h$ \cite{Bacchetta:2006tn}. In terms of
the TMD fragmentation functions we are interested in, this reads 
\begin{align}
  \label{eq:SIDIS_to_DIS_TMDlevel}
  \sum_h \int dz d^2p_{hT} z D_1^h(z,p_{hT}) & = \xi_2 = 1   \\
  \sum_h \int dz d^2p_{hT} E^h(z,p_{hT}) & = \xi_1 = \mj / \Lambda\ ,
\end{align}
where $D_1^h(z,p_{hT})$ is the twist-2 quark fragmentation function as a
function of the hadron's collinear momentum fraction $z$ and transverse
momentum $p_{hT}$, and $E^h(z,p_{hT})$ is a chiral-odd twist-3 function
defined in \cite{Bacchetta:2006tn}.

% As we shall see, an analogous formula for quark-gluon-quark correlators is not needed.
%The $\xi_1$ term is chiral-odd and therefore can appear in the inclusive cross section only coupled to the transversity function $h_1$. Therefore, for our analysis 
The relevant part of the semi-inclusive hadronic tensor for our analysis is 
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} 
    \sum_q e_q^2
    \bigg[ 2 \xbj g_T(\xbj) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1(\xbj) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots
\end{align}
For clarity, here we reintroduced the quark flavor $q$, $e_q$ being its electric charge.
The first term can be easily integrated with the help of the sum rule
\eqref{eq:SIDIS_to_DIS_TMDlevel}. To integrate the latter, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
make use of the sum rule \eqref{eq:SIDIS_to_DIS}: 
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \sum_h \int dz d^2p_{hT} \Big[ E^{q,h}(z,p_{hT}) - \frac{\mq}{\Lambda} z D_1^{q,h}(z,p_{hT}) \Big]
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda} \ .
\end{align}
This formula is the single most important result of this paper, and provides a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the integral would vanish.

Finally, the contraction of the hadronic tensor with the leptonic tensor leads
to the following well known result for the inclusive DIS cross section up to order $\Lambda/Q$~\cite{Bacchetta:2006tn}
\begin{align}
\frac{d\sigma}{d\xbj \, dy\, d\psi}
%&
=
\frac{2 \alpha^2}{\xbj y Q^2}\,
\frac{y^2}{2\,(1-\varepsilon)}\, 
\biggl\{
&F_{UU ,T} + \varepsilon F_{UU ,L}
+ S_\parallel \lambda_e\,
  \sqrt{1-\varepsilon^2}\; 
F_{LL}
%\nonumber 
%\\  &
+ |\bm{S}_\perp| \lambda_e\, \sqrt{2\,\varepsilon (1-\varepsilon)}\, 
  \cos\phi_S\, 
F_{LT}^{\cos \phi_S}
 \biggr\} \ ,
\label{e:crossdis}
\end{align}
where the structure functions on the right hand side correspond to
\begin{align}
F_{UU ,T} &= \xbj\,\sum_q e_q^2\,f_1^q(\xbj),
\\
F_{UU ,L} &= 0,
\\
F_{LL} &=\xbj\,\sum_q e_q^2\,g_1^q(\xbj),
\\
F_{UT}^{\sin \phi_S}&=0,
\label{e:FUTint}
\\
F_{LT}^{\cos \phi_S}&=-\xbj\,\sum_q e_q^2\, \frac{2\Lambda}{Q}\,
\biggl(\xbj  g_T^q(\xbj)
   + \frac{\mj -\mq}{\Lambda} \, h_{1}^q(\xbj) \biggr).
\label{e:FLTint}
\end{align}
\todo{Alessandro, please double check that $h_1$ appears naked in Eq.~\eqref{e:FLTint}, and that this implies dividing it by $x$ in Eq.~\eqref{e:g2}. It is fundamental for the conclusion about the ETL sum rule later on.\\}
The second term in the last structure function is a new result from our
analysis; it is not suppressed as an inverse power of $Q$ compared to the
standard term. On the nonperturbative
vacuum the jet mass is larger than the quark's, and this contributes a
nonnegligible term to the twist-3 part of the $g_2$ function, as we will
discuss in the next section.

\section{The $g_2$ structure function}

The new term in Eq.\eqref{e:FLTint} only appears in the $g_2$ structure function. \AAcom{But in the past we mentioned that it could also appear in some fancier structure function -- the one that also obays a WW-type relation. Or maybe, we were saying that $\xi_1$ might appear there, and the sentence as written here is correct.} Following the derivation in Ref.~\cite{Accardi:2009au}, one finds
\begin{align}
\label{e:g2}
  g_2(\xbj) = g_2^{WW}(\xbj) + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(\xbj) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(\xbj)}{\xbj} 
\Biggr) \ ,
\end{align}
where we defined $f^*(x) = -f(x) + \int_x^1\frac{dy}{y} f(y)$. The first four
terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the fifth is new. Note that even if
the relation is written for the sum of the quark weighted by their charge
squared, it can be considered valid also flavor by flavor. In fact, the steps
leading to such a decomposition are formulated at the correlator level.

The first term is also known as the Wandzura-Wilczeck function $g_2^{WW} =
-g_1^*$, and contains all the twist-2 chiral-even contributions to
the $g_2$ structure coming from quark-quark correlators. The second and third
terms contain all ``pure twist-3'' contributions, i.e., those coming from
quark-gluon-quark correlators. The fourth and fifth terms contain chiral-odd twist-2 contributions and depend on the
transversity parton distribution function, $h_1$. 
The fourth term is usually neglected for
light quarks since it is proportional to $\mq=O$(1 MeV). The last term,
new in our analysis, is again proportional to the transversity distribution but multiplied by a constant of $O$(100 MeV), and cannot be a priori neglected.

It is important to estimate the size of the various contributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{\rm tw3} & = \frac{1}{2}\,\sum_q e_q^2
    \biggl(
    \widetilde g_T^{q \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
\nonumber \\ 
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_q e_q^2 
 \frac{\mq}{\Lambda} (h_1^q/x)^\star(\xbj) \ ,
\\
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_q e_q^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(\xbj)}{\xbj} \ .
\nonumber 
\end{align} 
These terms are compared in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $\xbj$ and small $Q^2$ from Jefferson Lab, and simultaneously fits the higher-twist components of
$g_1$ and $g_2$ to the data. For the ``pure twist-3'' contribution,
$g_2^{\rm tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a recent light-front model calculation by Braun et al.~\cite{Braun:2011aw}; for
(modified) bag model calculations, see \cite{Jaffe:1990qh,Stratmann:1993aw}. To estimate the contributions
from quark ($g_2^{\text{quark}}$) and jet mass ($g_2^{\text{jet}}$) effects, that depend on chiral-odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

As one can see, in the proton case the pure twist-3 contribution is quite
smaller in magnitude, and opposite in sign, compared to the JAM15 fit. The quark-mass contribution, as expected, is essentially negligible. 
For what concerns the jet-mass contribution, the uncertainties due to the $h_1$
extraction are very large, especially at low $\xbj$. In addition, there is an
overall normalization uncertainty due to the choice of $\mj$, not shown in the
plot. In any case, the jet-mass contribution is strikingly large. If we assume
that the pure twist-3 contributions are of the order of the model calculation
by Braun et al., the breaking of the Wandura-Wilczek relation can be used to
constrain the extractions of the transversity distribution. This is inparticular true at low $\xbj$, where th wpure twist-3 term is expected to vanish. 
Moreover, it is quite clear that the gap
between the pure twist-3 $g_2^{\rm tw3}$ function and the JAM15 fit can be
explained by the new jet-mass contribution we discuss in this paper.

In the neutron case, the jet contribution is very negative at intermediate to
large values of $\xbj$. 
If one trusts the order of magnitude of the $g_2^{\rm tw3}$
calculation by Braun et al., one would conclude that the jet contribution
should not be that large. However, this contribution depends strongly on the $d$ quark's transversity, whose fit suffers from large systematic uncertainties and saturates the negative Soffer bound. Recent deata in $p+p$ collisions indicate, in fact, that $h_1^d$ might be less negative than in the Pavia15
fits~\cite{Radici:2016lam}.
Correspondingly the jet contribution to the proton at $\xbj \approx 0.1$
would become less positive, inproving as well the agreement with the JAM15
fit.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure function compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The unceratainty in the choice
$m_q=5$ GeV and $\mj=100$ GeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estimate was provided in the original reference). 
}
\end{center}
\end{figure}

\section{Moments of the $g_2$ strcuture function}

It is interesting to consider the moments of the non Wandzura-Wilczek contribution to $g_2$,
\begin{align}
  d_N \equiv (N+1) \int_0^1 x^N \bigg( g_2(x) - g_2^{WW}(x) \bigg) \ .
\end{align}
For a generic function $f$, let us define it's $N$-th moment as $f[N]=\int_0^1 dx\, x^{N} f(x)$. It is then straightforward to verify that $f^*[N] = N/(N+1) \times f[N]$ and  
\begin{align}
  d_N & = (N+1) g_2[N] + N g_1[N] \\
  & = \frac12 \sum_q e_q^2 \bigg( N \tilde g_T^q[N] + \hat g_T^q[N]
    + \frac{(N+1) \mj-\mq}{\Lambda} h_1^q[N] \bigg) \ .
\end{align}

The zero-th moment provides an interesting relationship between transversity
and the inclusive structure function $g_2$:
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \sum_q e_q^2 \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1^q(x) \ .
\end{align}
Here we used the fact $f^*[0]=0$ for any function $f$, which eliminates the contribution from the quark-mass and the twist-3 $\widetilde g_T$ terms; instead, $\widehat g_T^q[0]$ vanishes identically due to the symmetry
properties of the quark-gluon-quark correlators, or equivalently as a consequence of the Lorentz invariance of QCD interactions, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. The only surviving term on the right-hand side is the new jet contribution.
The sum rule \eqref{eq:BC}
generalizes the Burkhardt-Cottingham (BC) sum rule~\cite{Burkhardt:1970ti},
which states that  $\int_0^1 dx\, g_2(x) =0$. 
Here, we have shown that jet-mass corrections can directly violate the BC sum rule.
In fact, the possibility of a violation of the BC sum rule due to 
contributions from spin-flip processes was already mentioned in the original
derivation, but these do not show up in treatments that only
consider free field quark propagators for the struck quark
\cite{Jaffe:1996zw}.  Although we formulated \eqref{eq:BC} in terms of a sum over quark flavors in order to display a clear
connection to the structure function $g_2$, we stress that this is valid also
flavor by flavor, i.e., for each single flavor the only measurable nonzero contribution to the zeroth moment of the
structure function $g_2$ can come from the coupling between the jet mass and the transversity function\footnote{This conclusion is true even if the BC sum rule is broken by a $J = 0$ fixed pole with
non-polynomial residue \cite{Jaffe:1996zw}, since this would appear as a
$\delta(\xbj)$ contribution and would not be measurable.}.

One should notice that since $h_1$ is slowly driven to 0 by QCD evolution as $Q^2 \to \infty$, the BC sum rule is still satisfied at least
asymptotically.
At finite scales, however, the only way to preserve the validity of the
Burkhardt-Cottingham sum rule is if
%that is derived at the cross section level (under suitable assumptions on the
%high-energy behavior of te hspin-flip amplitude) rather than from an analysis
%of collinear correlators, as done here, 
%to be strictly valid, we obtain a
%strong constraint on the transversity function, 
\begin{align}
   \int dx\, \frac{1}{x} h_1^q(x) = 0 \ .
\label{eq:ABsumrule}
\end{align}
Interestingly, one can show that this constraint, if valid at any given scale
$Q_0$ is conserved through QCD evolution. However, we think that this
constrain cannot be satisfied in general, since it is explicitly broken in perturbative
QCD~\cite{Kundu:2001pk} as well as in model calculations (see, e.g., \cite{Schweitzer:2001sr,Wakamatsu:2007nc,Pasquini:2005dk,Cloet:2007em,Bacchetta:2008af,Bourrely:2010ng}). If, nonetheless,  we assume that the BC sum rule is broken by a {\em finite} amount, we
obtain that $h_1(x)/x$ must be integrable, implying a 
a bound on the small $x$ behavior of the transversity, 
\begin{align}
  h_1^q(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\label{eq:ABbound}
\end{align}
This bound can be very useful, \eg, in transversity fits, where the data at
small $x$ is, as yet, very limited, and in general for proper extrapolations
when measuring moments.

The first moment is the first one to display a contribution from the pure twist-3 part of $g_2$:
\begin{align}
  d_1 & = \frac12 \sum_q e_q^2 \bigg( 2 \tilde g_T^q[1] + \hat g_T^q[1]
    + \frac{2\mj-\mq}{\Lambda} h_1^q[0] \bigg)
\label{e:d1}
\end{align}
where $h_1^q[1] = \int_0^1 dx h_1^q(x)$ is the contribution of a quark $q$ to the target's tensor charge. 
The second moment is also interesting because the pure twist-3 part can be
related to quark-gluon-quark correlators, see \cite{Jaffe:1996zw}, and
interpreted as as the average color force experienced by the struck quark as
it exits the nucleon \cite{Burkardt:2012sd}: 
\begin{align}
  d_2 & = \frac12 \sum_q e_q^2 \bigg( 3 \tilde g_T^q[2] + \hat g_T^q[2]
    + \frac{3\mj-\mq}{\Lambda} h_1^q[1] \bigg) \ .
\end{align}
For experimental
measurements of $d_2$, see, e.g., Refs.~\cite{Anthony:2002hy,Slifer:2008xu,Solvignon:2013yun,Posik:2014usi,Flay:2016wie}.
For both of these moments, the transversity contribution is a background to the extraction of the pure twist-3 piece. Fortunately, it is a quantity that can be extracted from the lattice
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn} or fitted
\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa}.
Furthermore, the new sum rule \eqref{eq:ABsumrule} and the bound
\eqref{eq:ABbound} promise to improve future transversity fits. Funally, th $\mj$ mass parameter can as well be evaliated on the lattice, or also measured, \eg, in electron-positron collisions, as we will briefly discuss later. Therefore the pure twist-3 part can, in principle, be properly isolated and measured.

Finally, we note that the jet contribution also leads to an explicit breaking of the Efremov-Teriayev-Leader (ETL) sum rule \cite{Efremov:1996hd}, in which the pure twist-3 contribution to the first moment of $g_2-g_2^{WW}$ also disappears. To see this, let's define the valence contribution to a given structure function as $f^V=\frac12\sum_q e_q^2 (f^q-f{\bar q})$. Then, as shown in \cite{Efremov:1996hd}, $\tilde g_T^V[1]+\widehat g_T^V[1] = 0$, and from Eq.~\eqref{e:d1} we obtain
\begin{align}
  d_1^V = \frac12 \sum_q e_q^2 \frac{2\mj- \mq}{\Lambda} 
    \big(h_1^{q}[0]-h_1^{\bar q}[0]\big) \ .
\end{align}
Assuming $M_u \approx M_d \gg m_u , m_d$ and isospin symmetry of the proton and neutron, we can also see that 
\begin{align}
  d_1^V = \frac{M_{\text{light}}}{\Lambda} \delta_T(p) \ ,   
\end{align}
This gives an alternative way to experimentally access the proton tensor charge, $\delta_T(p) = \sum_q e_q^2 \big(h_1^{q}[0]-h_1^{\bar q}[0]\big)$, by measusing moments of the flavor separated $g_2$ structure function.
\AAcom{I tried to build an observable for $d_1^V$ such as $p-n$ or $p-\bar p$, but I am messing up a bit. We may try and discuss this at the Gordon.}
%, if the mass parameter $M_{\text{light}} \equiv M_u=M_d$ can be measured with sufficient accuracy, for example in positron-electron annihilation as we shall briefly discuss next.

\section{Measuring the jet PCFs}

To experimentally measure jet functions, and in particular the jet mass parameter $M_q$, a promising avenue is through inclusive single hadron production, $e^+ e^- \to h X$, and inclusive dihadron 
production from the same hemisphere, $e^+ e^- \to h h X$, see 
Fig.~\ref{fig:epl_emn_h}.
In single-hadron production, the fragmentation functions $D^h$
play the role of PDFs in DIS, and couple to the jet functions
in an analogous way. To access the spin-flip $J_1$ function one needs to detected a polarized hadron, such as a $\delta$ baryon. 
In double hadron production, the enlarged number of Dirac structures
of the dihadron fragmentation correlator allows one to access
the jet function in novel ways, and in particular to isolate the
contribution from the helicity-flip $J_1$ term in combination with the
chiral-odd fragmentation function $H_1^{\sphericalangle}$. While measurements in the asymptotic large $Q^2$ regime will provide access to the integral of the $J_1$ jet function, i.e., to the jet-mass parameter $\mj$, perhaps even more interesting is the possibility to experimentally measure at finite values of $Q^2$ the momentum dependence of the jet functions $J_1$ and $J_2$, that enter structure functions in a way analogous to Eq.~\eqref{eq:F2}. In other words, it becomes possible to experimentally access the quark's spectral function.

Studying and classifying all the possibilities offered by single and
double hadron production in electron-positron annihilation events will open up a rich phenomenology, which will in turn be needed to extract pure twist-3 matrix elements from the $g_2$ structure function, and more in general to perform precise jet mass corrections in DIS.

\begin{figure}[htb]
  \centering
  \parbox[b]{8cm}{\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_X}
  }
  \parbox[b]{8cm}{\vskip-0.5cm\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_h_X}
  }
  \caption{\small
	Single hadron {\it (left)} and double hadron {\it (right)}
	production in $e^+ e^-$ collisions at LO with jet and
	fragmentation correlators.
  }
  \label{fig:epl_emn_h}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

In this paper, we revisited the inclusive DIS analysis, including the
effects due to the production of a system of final state hadrons in the current direction, which we
conveniently referred to as a ``jet.'' We described this in terms of a jet
correlator that corresponds, up to twist-4 contributions, to the nonperturbative quark cut propagator, or, equivalently, to the quark's spectral function. 
%different from the standard perturbative propagator. 
We then carried out the analysis of 
%the jet correlator and 
the DIS cross section up to twist-3 level. 
The introduction of this jet correlator
leads to a difference in the expression of the structure function $g_2$ in
inclusive DIS with respect to the standard analysis: a new term appears, proportional to a jet mass
parameter and to the transversity distribution function. This new term
contributes to the violation of the Wandzura-Wilczek relation, in addition to
the standard pure twist-3 terms and quark mass corrections. Contrary to these
standard terms, however, the new jet mass correction does not necessarily integrate to zero and so violates also the
Burkhardt-Cottingham sum rule. This is yet another example of how surprising
and rich the phenomenology of polarized inclusive DIS can be. 
Detailed measurements of the $g_2$ structure
function can be used to constrain the jet mass parameter, the transversity
distribution function and the nucleon tensor charge, helping their extraction from other observables in, e.g.,
electron-positron annihilation and semi-inclusive DIS, respectively. 
The
knowledge of jet
mass parameter and transversity distribution will be eventually needed for a precise
extraction of pure twist-3 terms from the structure function $g_2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% ACKNOWLEDGMENTS %%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{acknowledgments}
This work was supported by the United States Department of Energy (DOE) contract No. DE-AC05-06OR23177,
under which Jefferson Science Associates, LLC operates Jefferson Lab, by the DOE contract DE-SC008791 and 
by the European Research Council (ERC) under the European Union's 
Horizon 2020 research and innovation programme (grant agreement No. 647981,
3DSPIN)
\end{acknowledgments}

\ \\ \AAcom{NOTE: the umlauts do not show up in the bibliography!! See, e.g., reference [5]. Need to fix this.}

\bibliographystyle{myrevtex}
\bibliography{jetbiblio}

\end{document}